---
title: "Simulate From EBE"
---

# Introduction
This blog post will show you how to use your model to do both Monte Carlo, 
population simulation and simulation from a specific set of empirical 
Bayes estimates (EBE) or specific individual parameters. The basic implementation
for this approach was pioneered by **Tim Waterhouse** at MetrumRG.

# Background

The situation is this: you've coded your population model in terms of 
`THETA`, `ETA`, some covariates and the like:

```{c, eval = FALSE}
[ param ] WT = 70 , BMI = 23

[ main ] 

double CL = THETA1 * pow(WT/70,0.75)    * exp(ETA(1));
double V  = THETA2 * pow(BMI/23,THETA3) * exp(ETA(2));
```

For this model, you feed in subject-specific `WT` and `BMI`, mrgsolve will 
simulate `ETA(1)` and `ETA(2)` and you will get a subject-specific `CL`
and `V2` that can be used to advance the system.

But what if you have a specific set of `CL` and `V` that came from your 
NONMEM model output that you want to used here instead?  For example, I have
this set of EBE that I want to use?


```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
set.seed(123)
par <- tibble(
  ID = c(1,2,3), 
  CL = runif(3, 0.5, 1.5), 
  V  = runif(3, 35, 67)
)
```

```{r}
par
```

How can I work these into these into the simulation with this model?

# Re-name the EBE

Instead of using `CL` and `V` in your data, rename them to `CLI` and `VI`

```{r}
par <- rename(par, CLI = CL, VI = V)
par
```

The specific names you choose isn't as important as naming them something 
other than what you used in the model.

# Add parameters to the model
Now you need to add these as parameters in the model so mrgsolve will pick them 
up when you pass them in on the data set

```{c, eval = FALSE}
[ param ] 
WT = 70 , BMI = 23
CLI = -1, VI = -1
```

Note we gave these parameters a special value: `-1`. This will be important 
for the next step.

# Conditional update of CL and V2

Now, after the population parameters are declared and defined, you can 
put a block of code that updates those variables only if `CLI` is greater than 
zero; that is, if the updated parameter is something other than that special / 
null value, we'll use that; otherwise use the random simulated value

```{c, eval = FALSE}
[ main ] 

double CL = THETA1 * pow(WT/70,0.75)    * exp(ETA(1));
double V  = THETA2 * pow(BMI/23,THETA3) * exp(ETA(2));

if(CLI > 0) {
  CL = CLI;
  V = VI;
}
```

# Include CLI and V2I in the data

Now, whenever we attach these specially named parameters to the data, we'll 
get a simulation from the EBE; otherwise, we'll get the simulation from 
the covariates and randomly drawn ETA.

# Full example

## Data

```{r, warning = FALSE, message = FALSE}
library(mrgsolve)
data <- expand.ev(amt = 100, ID = 1:3) %>% left_join(par)
data
```

## Model

```{r}
code <- '
[ param ] 
WT = 70 , BMI = 23, KA = 1
CLI = -1, VI = -1

[ param ] THETA1 = 1, THETA2 = 30, THETA3 = 0.21

[ main ] 

double CL = THETA1 * pow(WT/70,0.75)    * exp(ETA(1));
double V  = THETA2 * pow(BMI/23,THETA3) * exp(ETA(2));

if(CLI > 0) {
  CL = CLI;
  V = VI;
}

[ omega ] 0.2 0.8

[ pkmodel ] cmt = "depot,cent", depot = TRUE

[ error ] 
capture cp = cent/V;
'
mod <- mcode("ebe", code, capture = "CL,V")
```

## Simulate

```{r}
out <- mrgsim(mod, data) 
plot(out)
```

## Check

Verify that the `CL` and `V` in the output match the `CLI` and `VI` in the 
input

```{r}
data
count(as_tibble(out), ID, CL, V)
```

## Random simulation

Now, if we drop the EBE from the data, we get our random basic random simulation

```{r}
data2 <- mutate(
  data, 
  CLI = NULL, 
  V2I = NULL, 
  BMI = c(20, 23, 26), 
  WT = c(60, 70, 80)
)
mrgsim_df(mod, data2) %>% count(ID, CL, V)
mrgsim_df(mod, data2) %>% count(ID, CL, V)
```

